### 2016-10-18 ms/ms/ms.tex
misc bug fixes

Next, to improve efficiency, using NIMBLE we filter over latent states
to calculate model likelihoods in order to limit MCMC sampling to
top-level parameters (``filter''). We then use two additional
computational approaches to improve the efficiency of this model 1)
dynamic blocking of the parameters \citep[``filter + blocking'',
][]{turek2016efficient}, and 2) a custom MCMC specification where
slice samplers \citep{neal-03} are used for all parameters (``filter +
slice''). \flagged{other options we want to present?}

Slice samplers are a class of methods that sample from a target
distribution by using that fact that samples from any distribution can
be obtained by sampling uniformly from the area under that
distribution's probability density function curve.  The horizontal
coordinates of these uniform samples will provide samples from the
distribution of interest. Slice samplers have been shown to perform
well in situations where choosing a tuning parameter for a Metropolis
algorithm is difficult. When used to sample from the posterior
distribution of a univariate parameter, a slice sampler proceeds at
each iteration by first choosing a vertical coordinate sampled
uniformly between 0 and the height of the density curve at the
parameter value from the previous iteration.  Then, a horizontal
coordinate is chosen uniformly from the set of all possible parameter
values whose density is at least as great as the chosen vertical
coordinate.

@@

To improve the efficiency of this model, we first filtered over latent
states to calculate model likelihoods in order to limit MCMC sampling
to top-level parameters. We also vectorized all calculations that
would have require for loops in BUGS or JAGS. We then tried two
approaches to speed sampling of the top-level parameters 1) dynamic
blocking of the parameters \citep{turek2016efficient}, and 2) a custom
blocking scheme where the parameters of each species are blocked
together with adaptive random walk MCMC.

Random walk Metropolis algorithms \citep{metropolis1953equation}
sample from the posterior distribution of a parameter of interest by
first proposing a new parameter value from a normal proposal
distribution centered at the current value, and then deciding whether
to accept or reject the proposed value via the Metropolis ratio.
Although common, standard univariate or multivariate normal proposal
distributions can prove to be inefficient at sampling for models in
which parameters are highly correlated.  Adaptive random walk
Metropolis sampling \citep{haario98anadaptive} allows the correlation
structure of the posterior distribution of blocks of parameters to be
used to produce better proposals.  For an occupancy model where
parameters are highly correlated within species, this can provide much
more efficient sampling than a non-adaptive Metropolis algorithm.

@@

In the multi species occupancy models (Section \ref{sec:msss}),
practitioners are often interested in determining whether a model
including a species random effect for explanatory variables is a
better fit than a model without the random effect. We implemented a
cross-validation procudure for this model where the detection data for
species is left out, the model refitted, and the fitted model used to
predict the occurence of that species. The predictive error of the
model included a random effect of species is then compared to a model
where no species random effects were included.

>>>>

Next, to improve efficiency, using NIMBLE we filter over latent states
to calculate model likelihoods in order to limit MCMC sampling to
top-level parameters (``filter''). \flagged{Do we want to write out
  the likelihood?)} We then use two additional computational
approaches to improve the efficiency of this model 1) dynamic blocking
of the parameters \citep[``filter + autoblocking'',
][]{turek2016efficient}, and 2) a custom MCMC specification where
slice samplers \citep{neal-03} are used for all parameters (``filter +
slice''). Slice samplers are a class of methods that sample from a
target distribution by using that fact that samples from any
distribution can be obtained by sampling uniformly from the area under
that distribution's probability density function curve.  The
horizontal coordinates of these uniform samples will provide samples
from the distribution of interest. Slice samplers have been shown to
perform well in situations where choosing a tuning parameter for a
Metropolis algorithm is difficult. When used to sample from the
posterior distribution of a univariate parameter, a slice sampler
proceeds at each iteration by first choosing a vertical coordinate
sampled uniformly between 0 and the height of the density curve at the
parameter value from the previous iteration.  Then, a horizontal
coordinate is chosen uniformly from the set of all possible parameter
values whose density is at least as great as the chosen vertical
coordinate.  \flagged{other options we want to present?}

@@

We first fit the using the default settings for JAGS and NIMBLE where
all model parameters and latent states undergo MCMC sampling
(``NIMBLE-latent'' and ``JAGS-latent'', respectively). We use
uninformative priors $\mathrm{Norm}(0,1000)$ for the means of the
distributions of the hyperparameters and $\mathrm{Unif}(0,100)$ the
variances.

To improve the efficiency of this model, we first filtered over latent
states to calculate model likelihoods in order to limit MCMC sampling
to top-level parameters (``Filter''). We also vectorized all
calculations that would have require for loops in JAGS. \flagged{Do we
  want to write out the likelihood?)} We then applied two approaches
to speed sampling of the top-level parameters 1) dynamic blocking of
the parameters \citep[``Filter + autoblocking'',
][]{turek2016efficient}, and 2) a custom blocking scheme where the
parameters of each species are blocked together with adaptive random
walk MCMC (``Filter + species blocking''). Random walk Metropolis
algorithms \citep{metropolis1953equation} sample from the posterior
distribution of a parameter of interest by first proposing a new
parameter value from a normal proposal distribution centered at the
current value, and then deciding whether to accept or reject the
proposed value via the Metropolis ratio.  Although common, standard
univariate or multivariate normal proposal distributions can prove to
be inefficient at sampling for models in which parameters are highly
correlated.  Adaptive random walk Metropolis sampling
\citep{haario98anadaptive} allows the correlation structure of the
posterior distribution of blocks of parameters to be used to produce
better proposals.  For an occupancy model where parameters are highly
correlated within species, this can provide much more efficient
sampling than a non-adaptive Metropolis algorithm.

@@

In multi-species occupancy models like the model we explored in
Section \ref{sec:msss}), practitioners are often interested in
determining whether a model including a species random effect for
explanatory variables is a better fit than a model without the random
effect. We implemented a cross-validation procedure for this model
where the detection data for species is left out, the model refitted,
and the fitted model used to predict the occurence of that
species. The predictive error of the model included a random effect of
species is then compared to a model where no species random effects
were included.
